Performance comparison of Dask and Apache Spark on HPC systems

### 2021-09-03 latex/paper.tex
Minor edits

\begin{abstract}
	% TODO
	The joint increase in data resolution, data sharing and
	study sizes motivates the adoption of Big Data strategies in a number of scientific disciplines.
	However, while several options are available, no particular Big Data engine emerged.
	In this paper, we compare the runtime performance of two popular Big Data engines with
	Python APIs, Apache Spark, and Dask, in processing neuroimaging pipelines.
	Our experiments use three synthetic data-intensive applications to process
	the \SI{606}{\gibi\byte} BigBrain image and an actual pipeline to process data
	from thousands of anatomical images.
	We benchmark these applications on a dedicated HPC cluster running the Lustre file system while using varying
	combinations of number of nodes, file size, and task duration.
	Our results show that although there are slight differences between Dask and Spark,
	the performance of the engines is comparable for data-intensive applications.
	However, Spark requires more memory than Dask, which can lead to 
	slower runtime depending on configuration and infrastructure.
	In general, the limiting factor was the data transfer time.
	While both engines are suitable for neuroimaging, more efforts need to be put
	to reduce the data transfer time and the memory footprint of applications.

@@

While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructures.
This paper focuses on the performance of Big Data engines in HPCs environments.
We installed the Slurm scheduler and Lustre file system~\cite{braam2019lustre} on a dedicated cluster to mimic a typical HPC system.
Lustre is a powerful and scalable file system for HPC environments offering parallel file accesses.

@@

\subsection{Histogram}
Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} application when varying the number of nodes.
Overall, Spark has up to 5\% lower total execution time, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
This is unexpected as the code base for both engines is the same, with only the calling APIs differing.
It is most likely due to the use of Dask Bags, which are slower than the other Dask APIs.
As with previous experiments, the overhead for Spark is less and scales better, although compensated by a longer total I/O time.
\begin{figure}[!t]
	\centering
	\includegraphics[clip,width=\columnwidth]{figures/stacked_histogram_worker.jpg}
	\caption{Histogram: varying nodes -- 5000 files, 5 iterations, \SI{1}{\second} delay}
	\label{fig:histogram_worker}
\end{figure}

@@

\subsection{Kmeans}
Figure \ref{fig:kmeans_worker} depicts the total execution time for the \textit{Kmeans} application when varying the number of workers.
We observe that Dask is multifold faster than Spark, resulting in a $1.79\times$ to $2.71\times$ lower makespan.
Specifically, Spark has much longer overhead, \textit{get\_labels} and read time than Dask.
This difference arises from Spark workers restarting due to lack of executor memory, leading to recomputation of tasks.

@@

\subsection{BIDS App Example}
Figure \ref{fig:bids} shows the total execution time for the \textit{BIDS App Example} application when varying the number of nodes.
We observe that Dask has slightly lower overhead, thank Spark, although not significant.
However, the makespan for Dask is between 7.5\% to 14.5\% lower than Spark.
The main difference comes from an increased amount of stagger tasks for Spark, thus increasing the overhead of the application.
Moreover, although minimal, a shorter cluster deployment time for Dask reinforces the difference in overhead time.
As with other experiments, the I/O from Lustre affects the scaling of both engines as we increase the number of nodes.
\begin{figure}[!t]
	\centering
	\includegraphics[clip,width=\columnwidth]{figures/stacked_bids.jpg}
	\caption{BIDS App Example: varying nodes}
	\label{fig:bids}
\end{figure}

\section{Discussion}
\subsection{I/O bottleneck}
We used Lustre as shared file system as it is a standard solution in HPC systems.
The performance of Lustre degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
In a previous study \cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
In this paper, Lustre shows some scaling when increasing the number of nodes, although not linear.
This suggests that increasing the parallelism level of the file system might help alleviate the I/O bottleneck observed.

>>>>

\begin{abstract}
	% TODO
	The general increase in data size and data sharing motivates the adoption of Big Data strategies in a number of scientific disciplines.
	However, while several options are available, no particular guidelines exist for the selection of a Big Data engine.
	In this paper, we compare the runtime performance of two popular Big Data engines with
	Python APIs, Apache Spark, and Dask, in processing neuroimaging pipelines.
	Our experiments use three synthetic data-intensive applications to process
	the \SI{606}{\gibi\byte} BigBrain image and an actual pipeline to process data
	from thousands of anatomical images.
	We benchmark these applications on a dedicated HPC cluster running the Lustre file system while using varying
	combinations of number of nodes, file size, and task duration.
	Our results show that although there are slight differences between Dask and Spark,
	the performance of the engines is comparable for data-intensive applications.
	However, Spark requires more memory than Dask, which can lead to 
	slower runtime depending on configuration and infrastructure.
	In general, the limiting factor was the data transfer time.
	While both engines are suitable for neuroimaging, more efforts need to be put
	to reduce the data transfer time and the memory footprint of applications.

@@

While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructures.
This paper focuses on the performance of Big Data engines in HPCs environments.
We installed the Slurm scheduler and Lustre file system~\cite{braam2019lustre} on a dedicated cluster to mimic a typical HPC system.
Lustre is a popular file system for HPC environments offering parallel file accesses.

@@

\subsection{Histogram}
Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} application when varying the number of nodes.
Overall, Spark's total execution time is up to 5\% lower than Dask's, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
This is unexpected as the code base for both engines is the same, with only the calling APIs differing.
It is most likely due to the use of Dask Bags, which are slower than the other Dask APIs.
As with previous experiments, the overhead for Spark is less and scales better, although compensated by a longer total I/O time.
\begin{figure}[!t]
	\centering
	\includegraphics[clip,width=\columnwidth]{figures/stacked_histogram_worker.jpg}
	\caption{Histogram: varying nodes -- 5000 files, 5 iterations, \SI{1}{\second} delay}
	\label{fig:histogram_worker}
\end{figure}

@@

\subsection{Kmeans}
Figure \ref{fig:kmeans_worker} depicts the total execution time for the \textit{Kmeans} application when varying the number of workers.
We observe that Dask is multifold faster than Spark, resulting in a $1.79\times$ to $2.71\times$ speed-up ratio on the makespan.
Specifically, Spark has much longer overhead, \textit{get\_labels} and read time than Dask.
This difference arises from Spark workers restarting due to lack of executor memory, leading to recomputation of tasks.

@@

\subsection{BIDS App Example}
Figure \ref{fig:bids} shows the total execution time for the \textit{BIDS App Example} application when varying the number of nodes.
We observe that Dask has slightly lower overhead, thank Spark, although not significant.
However, the makespan for Dask is between 7.5\% to 14.5\% lower than for Spark.
The main difference comes from an increased amount of stagger tasks for Spark, thus increasing the overhead of the application.
Moreover, although minimal, a shorter cluster deployment time for Dask reinforces the difference in overhead time.
As with other experiments, the I/O from Lustre affects the scaling of both engines as we increase the number of nodes.
\begin{figure}[!t]
	\centering
	\includegraphics[clip,width=\columnwidth]{figures/stacked_bids.jpg}
	\caption{BIDS App Example: varying nodes}
	\label{fig:bids}
\end{figure}

\section{Discussion}
\subsection{I/O bottleneck}
We used Lustre as shared file system as it is a standard solution in HPC systems.
The performance of our Lustre installation degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
In a previous study \cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
In this paper, Lustre shows some scaling when increasing the number of nodes, although not linear.
This suggests that increasing the parallelism level of the file system, including through potential configuration optimizations, might help alleviate the I/O bottleneck observed.
